---
title: "7CS034 DATA SCIENCE ASSESSMENT 2406632"
author: "Martins Timilehin"
format:
  html:
    code-fold: true
    code-link: true
    toc: true
    toc-location: left
    toc-expand: true
    number-sections: true
bibliography: references.bib
editor: visual
---

```{=html}
<style>
h1 {
  font-size: 20pt;
}
</style>
```
```{=html}
<style>
h2 {
  color: #424858;
  font-size: 16pt;
}
</style>
```
# Loading R packages

I used tidyverse documentation [@tidyverse2019] for data cleaning and R Programming code for exploration and visualization [@rsoftware] and Quarto [@quarto2022] documentations for markdown.

```{r}
#| label: load-packages
#| include: true
#| eval: true
#| results: hide
library(ggplot2)
library(patchwork)
library(tidyverse)
```

# Reading dataset into RStudio

```{r}
data <- read.csv("C:/Users/2406632/Downloads/data.csv")
head(data)
```

# Data preprocessing & feature engineering

## Handling missing values

Missing values spotted in the data are below: Mar_Sat=6,144 \| INC=6,173 \| Highest.Ed=1,123

```{r}
clean_data <- data[complete.cases(data), ]
```

## Categorize Age data into Age_Group

```{r}
age_breaks <- c(-Inf, 10, 20, 30, 40, 50, 60, Inf)
age_labels <- c("<10", "10-19", "20-29", "30-39", "40-49", "50-59", "60+")
clean_data$Age_Group <- cut(clean_data$Age, breaks = age_breaks, labels = age_labels, right = FALSE)
```

## Renaming columns

```{r}
new_column_names <- c("Marital_Status", "Income", "Gender", "all_rooms_behind_door", "Ethnicity", "Highest_education")
colnames(clean_data)[c(4, 5, 6, 7, 8, 9)] <- new_column_names
```

## Renaming column values of gender

0:Female \| 1:Male

```{r}
clean_data$Gender <- ifelse(clean_data$Gender == 0, "Female", "Male")
clean_data$all_rooms_behind_door <- ifelse(clean_data$all_rooms_behind_door == 0, "Yes", "No")
```

## Taking a peek at the Cleaned data

```{r}
head(clean_data)
```

# Data Visualization

## Demographic Overview - Population Pyramid
The graph shows a majority in the 60+ age group with balanced gender distribution, while younger groups are less represented.

```{r}
ggplot(clean_data, aes(x = Age_Group, fill = Gender)) +
  geom_bar(stat = "count", position = "dodge") +
  coord_flip() +
  labs(title = "Demographic Overview - Age Distribution",
       x = "Count",
       y = "Age Group")
```

## Marital Status Composition - Pie Chart
The pie chart shows marital status distribution, with "Married" and "Never Married" being the largest categories, while "Divorced," "Widowed," and "Separated" have smaller proportions.

```{r}
ggplot(clean_data, aes(x = "", fill = Marital_Status)) +
  geom_bar(width = 1, stat = "count") +
  coord_polar(theta = "y") +
  labs(title = "Marital Status Composition",
       fill = "Marital Status") +
  theme_void()
```

## Income Distribution - Histogram
The histogram shows a right-skewed income distribution, with most individuals earning low annual incomes, while higher income levels are represented by fewer individuals.

```{r}
ggplot(clean_data, aes(x = Income)) +
  geom_histogram(binwidth = 5000, fill = "skyblue", color = "black") +
  labs(title = "Income Distribution",
       x = "Annual Income (in pounds)",
       y = "Count")
```

## Privacy Preferences - Stacked Bar Chart
The bar chart shows that the majority prefer privacy ("Yes"), with significantly fewer selecting "No," indicating strong privacy concerns among the surveyed individuals.

```{r}
ggplot(clean_data, aes(x = factor(all_rooms_behind_door), fill = factor(all_rooms_behind_door))) +
  geom_bar() +
  labs(title = "Privacy Preferences",
       x = "Response to Question H8",
       y = "Count")
```

## Education distribution
The bar chart shows education levels, with "Secondary School" and "Some HE" being the most common, while "Masters or higher" has the fewest individuals, indicating varied educational attainment in the dataset.

```{r}
plot5 <- clean_data %>%
  count(Highest_education) %>%
  ggplot(aes(x = Highest_education, y = n)) +
  geom_bar(stat = "identity") +
  ggtitle("Education Distribution") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot5
```

## Age Distribution
The bar chart shows the age group distribution, with the 60+ group having the highest count, while the 10-19 group has the lowest, indicating an older-skewed population.

```{r}
plot6 <- clean_data %>%
  count(Age_Group) %>%
  ggplot(aes(x = Age_Group, y = n)) +
  geom_bar(stat = "identity") +
  ggtitle("Age Group Distribution") +
  theme(axis.text.x = element_text(size = 15))

plot6
```

## Education levels by gender
The bar chart shows education levels by gender, with "Secondary School" and "Some HE" being the most common for both genders, while "Masters or higher" is the least common.

```{r}
plot7 <- ggplot(clean_data, aes(x = Highest_education, fill = Gender)) +
  geom_bar(position = "dodge", show.legend = TRUE) +
  labs(title = "Distribution of Education Levels by Gender")

plot7  + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Income distribution by family structure and age
The scatter plot shows income rising with age, with "Married" individuals having higher incomes across most age groups.

```{r}
ggplot(clean_data, aes(x = Age, y = Income, color = Marital_Status)) +
  geom_point(size = 3) +  # Adjust the size as needed
  theme(
    plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")  # Adjust the margins as needed
  ) +
  labs(title = "Scatter Plot of Age vs Income by Marital Status")
```

# Conclusion
The data predominantly features older individuals (60+), with "Married" and "Never Married" as the leading marital statuses. Income increases with age, particularly among married individuals. Privacy is a significant concern.

# Recommendation
Focus on marketing to older, high-income individuals, particularly those who are married. Develop privacy-focused products to align with user preferences. Additionally, expand data collection to include more representation from younger age groups for a more balanced analysis.

```{references}
```
